Intentional release of native species undermines ecological stability

### 2022-12-23 rmd/manuscript.Rmd
edit: correct cross-reference

$f_R~(\ge 0)$ is the density-independent survival of captive-bred individuals relative to wild individuals. Equation \@ref(eq:ricker) can be reorganized to:

@@

where $\xi_{1,s}$ is the site-specific constant at site $s$, $\xi_{2,s}$ the parameter characterizing an autoregressive (AR) process, and $\varepsilon_{s,t}^{\text{state}}$ the process error that follows a normal distribution as $\varepsilon_{s,t}^{\text{state}} \sim \text{Normal}(0, \sigma_{\text{state},s}^2)$. The parameters $\xi_{1,s}$ and $\xi_{2,s}$ were drawn from a multivariate normal distribution as $\pmb{\xi} \sim \text{MVN}(\pmb{\theta_{\xi}}, \pmb{\Omega_{\xi}})$, where $\pmb{\xi}$ is the matrix of $\xi_{1,s}$ and $\xi_{2,s}$. Equation \@ref(eq:state) is structurally equivalent to a $Q$-order AR model; however, we expressed the part of AR parameters as a geometric-series $\tau^q$ to reduce the number of parameters (our model has two AR parameters $\xi_{2,s}$ and $\tau$ regardless of order $Q$). The parameter $\tau$ controls how fast the density influence attenuates towards the past [@takagiTimescaleDependencyHost2015]. In this study, we set $Q=3$ given the life span of the study species (typically 1 -- 3 years). We used median estimates of fish density $n_{s,t}$ to calculate the temporal CV, mean ($\mu$), and SD ($\sigma$) for each site. For a whole community, we summed the densities of enhanced and unenhanced species. We summarized the reconstructed community dynamics in **Figures S**11 -- **S**13.

@@

In Equation \@ref(eq:effr), $\eta_{w,s'}$ represents the effective proportion of total release affecting the community at site $s'$ within watershed $w$. The summation of $\eta_{w,s'}$ over $S_w+1$ is assumed to be unity because a non-negligible portion of release may escape into unsurveyed areas. Since the values of $\eta_{w,s'}$ are not available in our dataset, we estimated them through a stochastic search while fitting the regression model to the data within a Bayesian framework (see **Model fitting**). Other site-level predictors include upstream watershed area (log-transformed), air temperature, precipitation, forest land use, and the number of observation years. The number of observation years was included as a control factor given the potential influence on the estimates of temporal community variability. Urban and agricultural land use was omitted because of either a limited value range (**Figure S**14) or a strong correlation with forest land use (**Figure S**15). The watershed-specific intercept was related to a watershed-level predictor as:

@@

Since the CV is an additive function of mean $\mu$ and SD $\sigma$ in a log scale, we can derive the effects of environmental factors on the CV from Equation \@ref(eq:model):

@@

Equation \@ref(eq:ricker-ssm) is equivalent to Equation \@ref(eq:ricker0) except the scale of competition coefficients $\alpha'_{ij}$ ($\frac{\alpha'_{ij}}{\alpha'_{ii}}=\alpha_{ij}$, where $\alpha'_{ii}=\frac{r_i}{K_i}$). We used sparse priors for $\alpha'_{ij}$ to reduce the risk of over-fitting [@mutshindaWhatDrivesCommunity2009]:

@@

In Equation \@ref(eq:sparse), the latent variable $z_{ij}$ is drawn from a Bernoulli distribution and controls the prior for $\alpha'_{ij}$. We used $c_1 = 1$ and $c_0 = 0.01$ so that $\alpha'_{ij}$ is freely estimated when $z_{ij}$ equals one ($\sigma^2_{\alpha} = 1$). Otherwise, the probability density of $\alpha'_{ij}$ would be concentrated around zero ($\sigma^2_{\alpha} = 0.01$), effectively omitting the parameter from the model [@mutshindaWhatDrivesCommunity2009]. We assigned different "inclusion" probabilities to diagonal (intraspecific competition) and off-diagonal elements (interspecific competition) as $z_{ii} \sim \text{Bernoulli}(p_{\alpha}^{\text{intra}})$ and $z_{ij,i \ne j} \sim \text{Bernoulli}(p_{\alpha}^{\text{inter}})$.

>>>>

$f_R~(\ge 0)$ is the density-independent survival of captive-bred individuals relative to wild individuals. Equation (2) can be reorganized to:

@@

where $\xi_{1,s}$ is the site-specific constant at site $s$, $\xi_{2,s}$ the parameter characterizing an autoregressive (AR) process, and $\varepsilon_{s,t}^{\text{state}}$ the process error that follows a normal distribution as $\varepsilon_{s,t}^{\text{state}} \sim \text{Normal}(0, \sigma_{\text{state},s}^2)$. The parameters $\xi_{1,s}$ and $\xi_{2,s}$ were drawn from a multivariate normal distribution as $\pmb{\xi} \sim \text{MVN}(\pmb{\theta_{\xi}}, \pmb{\Omega_{\xi}})$, where $\pmb{\xi}$ is the matrix of $\xi_{1,s}$ and $\xi_{2,s}$. Equation (7) is structurally equivalent to a $Q$-order AR model; however, we expressed the part of AR parameters as a geometric-series $\tau^q$ to reduce the number of parameters (our model has two AR parameters $\xi_{2,s}$ and $\tau$ regardless of order $Q$). The parameter $\tau$ controls how fast the density influence attenuates towards the past [@takagiTimescaleDependencyHost2015]. In this study, we set $Q=3$ given the life span of the study species (typically 1 -- 3 years). We used median estimates of fish density $n_{s,t}$ to calculate the temporal CV, mean ($\mu$), and SD ($\sigma$) for each site. For a whole community, we summed the densities of enhanced and unenhanced species. We summarized the reconstructed community dynamics in **Figures S**11 -- **S**13.

@@

In Equation (10), $\eta_{w,s'}$ represents the effective proportion of total release affecting the community at site $s'$ within watershed $w$. The summation of $\eta_{w,s'}$ over $S_w+1$ is assumed to be unity because a non-negligible portion of release may escape into unsurveyed areas. Since the values of $\eta_{w,s'}$ are not available in our dataset, we estimated them through a stochastic search while fitting the regression model to the data within a Bayesian framework (see **Model fitting**). Other site-level predictors include upstream watershed area (log-transformed), air temperature, precipitation, forest land use, and the number of observation years. The number of observation years was included as a control factor given the potential influence on the estimates of temporal community variability. Urban and agricultural land use was omitted because of either a limited value range (**Figure S**14) or a strong correlation with forest land use (**Figure S**15). The watershed-specific intercept was related to a watershed-level predictor as:

@@

Since the CV is an additive function of mean $\mu$ and SD $\sigma$ in a log scale, we can derive the effects of environmental factors on the CV from Equation (9):

@@

Equation (15) is equivalent to Equation (1) except the scale of competition coefficients $\alpha'_{ij}$ ($\frac{\alpha'_{ij}}{\alpha'_{ii}}=\alpha_{ij}$, where $\alpha'_{ii}=\frac{r_i}{K_i}$). We used sparse priors for $\alpha'_{ij}$ to reduce the risk of over-fitting [@mutshindaWhatDrivesCommunity2009]:

@@

In Equation (16), the latent variable $z_{ij}$ is drawn from a Bernoulli distribution and controls the prior for $\alpha'_{ij}$. We used $c_1 = 1$ and $c_0 = 0.01$ so that $\alpha'_{ij}$ is freely estimated when $z_{ij}$ equals one ($\sigma^2_{\alpha} = 1$). Otherwise, the probability density of $\alpha'_{ij}$ would be concentrated around zero ($\sigma^2_{\alpha} = 0.01$), effectively omitting the parameter from the model [@mutshindaWhatDrivesCommunity2009]. We assigned different "inclusion" probabilities to diagonal (intraspecific competition) and off-diagonal elements (interspecific competition) as $z_{ii} \sim \text{Bernoulli}(p_{\alpha}^{\text{intra}})$ and $z_{ij,i \ne j} \sim \text{Bernoulli}(p_{\alpha}^{\text{inter}})$.
